---
title: "Electronic Trading Liquidity Management"
author: 
  - Kgothatso Baloyi 
  - Singalakha Menziwa
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: darkly
    highlight: tango
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
wd_global <- "C:/Users/r5478278/OneDrive - FRG/Documents/Grad Rotation info/Singalakha/5. Electronic Liquidity/Shiny/IntraDay_Analytics"
knitr::opts_knit$set(root.dir = wd_global)
setwd(wd_global)

# Functions
cache_data_loc <- "cache/run_time"

# Library
library(plotly)
library(htmlwidgets)
library(htmltools)
library(xml2)
library(dplyr)
```

# eFX Daily colour
## US 

(*Jan-27*) The FED is expected to hold its key interest rate steady this week[29th of Jan]. After three consecutive rate cuts since September, the Fed is now in a "wait and see" mode, looking for more data indicating that inflation is cooling. The current rate is in the range of 4.25%[LB] - 4.50%[UB].

(*Jan-28*) The US dollar strengthened today against all Group-of-10 currencies after Trump announced plans for significantly larger universal tariffs than the 2.5% proposed by Treasury Secretary Scott Bessent. Trump indicated upcoming tariffs on foreign-produced semiconductors, pharmaceuticals, and certain metals to encourage domestic manufacturing. 
Consequently, copper prices fell amid concerns of potential trade wars due to the planned tariffs on metals like aluminum and steel.

(*Jan-29*) Gold prices could reach or approach record highs if the FED announces a dovish pause today. This outcome seems likely due to various factors, including ongoing trade tensions. Currently, gold prices are just above $2,763 per ounce.

## ECB

(*Jan-27*) The market expects the ECB to cut its benchmark deposit rate by 25bps to 2.75% on Thursday[30th of Jan]. This would be the fourth consecutive rate cut as the ECB aims to support the eurozone economy and manage inflation.

## South Africa

(*Jan-27*) The SARB is expected to cut interest rates by 25bps this week[30th of Jan] following last weeks softer-than-expected inflation print. Currently repo is at 7.75%.

(*Jan-28*) The market has stabilized, and the rand has improved following the DA reaffirmation of their commitment to the GNU. This comes after weekend reports suggested instability within the GNU due to the President signing a new law facilitating state land expropriation. 

(*Jan-29*) ZAR still under pressure until SARB announcement.

## USDZAR levels

```{r, echo=FALSE,  fig.width=12, fig.height=7}
curr_pair = "USDZAR"
combined_volume_plot <- readRDS(file.path(cache_data_loc, paste0("combined_vol_price_", curr_pair, ".rds")))

# Display the plot
combined_volume_plot
```

- At 1 PM (*Jan-27*), the ZAR surged past our resistance level of 18.60, driven by idiosyncratic risks linked to the expropriation bill. This upward movement is notable given the concurrent decline in the dollar index and gains in gold prices, both of which typically support the ZAR.
  - Manqoba [RMB's Research Analyst] notes that the weakening correlation between gold and the South African Rand (ZAR) is expected, given the significant decline in gold mining output. As this downward trend in production continues, the correlation and its impact on terms of trade are likely to diminish further.
- The ZAR regains some ground (*Jan-28*) after the DA confirmed its commitment to the GNU. The ZAR is now trading below 18.70, signaling a positive market response to this news.
- When we broke the 18.60 lvl at 1PM (*Jan-27*), the implied topside was 18.87 and we later traded a high of 18.86 on (*Jan-28*). Thus our updated usdzar range is 18.87 [up 27cents] - 18.50 [up 10cents]. Notice the topside is up 27cents from prior topside range while the bottom is up just 10cents, this is indicative of the topside pressure on the ZAR.
- Based on charts, one can put topside at 19.00 for ZAR.

## Key events this week:
- Fed rate decision followed by news conference by Chair Jerome Powell, Wednesday
- Canada rate decision, Wednesday.
- Tesla, Microsoft, Meta, ASML earnings, Wednesday [At the back of DeepSeek]
- Eurozone ECB rate decision, consumer confidence, unemployment, GDP, Thursday
US GDP, jobless claims, Thursday
- Apple, Thursday [At the back of DeepSeek]
- ECB rate decision followed by news conference by President Christine Lagarde, Thursday
- US personal income & spending, PCE inflation, employment cost index, Friday

# Africa
*By sizwe Mfayela - Institutional Sales Specialist*

(*Jan-29*)

- Botswana: 
  - President Duma Boko reached a deal with De Beers for diamond extraction and sales. The terms of the agreement are set to be announced soon and the deal will help the country focus on promoting it’s natural diamonds.

- Egypt: 
  - Egypt issued \$2bio in Eurobonds yesterday, for the first time in two-years. The issuance was split between \$1.25bio in 5Y bonds at 8.625% and \$750mio in 8Y bonds at 9.45%.

- Mozambique: 
  - Mozambique missed its 2024 revenue target by more than 10% following the post-election protests and unrest, which cost the economy c.\$664mio in lost revenues. The state collected MZN 344.8bio (\$5.4bio) in revenues.

  - President Daniel Chapo met with Total Energies CEO, Patrick Pouyanne to discuss the company’s projects in Mozambique. Pouyanne confirmed Total Energies’ commitment to resume a $20bio LNG project that has been suspended since 2021. President Chapo emphasised the government is making efforts to ensure the necessary stability for the project’s implementation.

- Nigeria: 
  - Nigeria is looking to raise $15bio from private investors to add to it’s power grid and provide electricity to 86 million citizens currently without electricity. Nigeria aims to increase its renewable energy power sources from 22% to 50% in the next 5-years. The country will offer higher tariffs to investors while providing subsidies to households to shield them from the increased electricity costs.

- Tanzania: 
  - President Samia Suluhu Hassan said Tanzania needs \$13bio in energy investment by 2030 as the country aims to diversify its energy mix. Hassan said the goal is to add 2,463MW of electricity from solar, wind, geothermal and natural gas sources by 2030. \$5bio of the projects’ investment is set to come from the private sector and the government will foot the rest of the bill.

- Zimbabwe: 
  - Jan MoM Zig Inflation accelerated to 10.50% vs Dec 3.7%. ZiG inflation also rose to 11.5% in USD terms vs Nov 0.6%.
  
- Eurobonds:  
  - SSA flows were on the thinner side yesterday but were two-way in general. The market opened bid to start, with KENINT outperforming on the back of locals adding to the risk post the small debt for nature swap headline, but soon turned softer on the back of the jittery US equities market open. ANGOL outperformed NGERIA, with hedge fund buyers adding to the risk in the afternoon to close +0.125-0.25pts.

# Options

*By Thuto Mukena - Institutional Sales Specialist*

(*Jan-29*) Yesterday’s session brought some late recovery for the ZAR following earlier losses. Initially, the Rand weakened to R18.8657/\$ amid no new developments on Deepseek and the signing of the Expropriation Bill into law. These movements nudged 1-week realized volatility higher to trade close to the implied vol market. The unexpected nature of this week’s developments, particularly around the signing of the Expropriation Bill, helps explain why 1-week implied vol seems to have underpriced the risk premium for this week. Despite the narrowed 1-week volatility risk premium, options participants are still holding onto some event risk premium. The 1-week implied vol tenor is trading at a 0.54 vol premium over the 1-month tenor, reflecting the market's caution over potential near-term surprises.

Implied volatility across both G10 and EM currencies edged lower in yesterday’s session as markets positioned ahead of the FOMC rate decision. USD/JPY 1-week implied vol led the decline, dropping 83bps from the open, while AUD/USD 1-week implied vol followed suit, easing by 18bps from opening levels. On the EM front, USD/TRY 1 week implied vol saw the most significant move, plunging by 538bps. USD/MXN 1-week implied vol also edged lower, dropping by 24bps from opening levels, as market anxiousness sees some cooling ahead of the rate decision. The FOMC first in line for this week, Fed Funds pricing low probability of a rate cut.

---
